Drivers of tree regeneration remain challenging to be captured by models of forest dynamics

### 2022-11-19 manuscript/manuscriptRW.Rmd
lots of writing in the manuscript and bullet poits in discussion

---
title: "Capturing regeneration in models of forest dynamics"
author:
- Olalla Díaz-Yáñez, Forest Ecology ETH Zurich
- Yannek Käber, Forest Ecology ETH Zurich
- Mats Mahnken, PIK Potsdam
- Samuel Fischer, UFZ
- Rico Fischer, UFZ
- Friedrich Bohn, UFZ
- Xavier Morin, cefe cnrs
- Manfred Lexer, BOKU
- Christian Hochauer, BOKU
- Katarina Merganicova, forim
- Tobias Mette, LWF Bayern
- Werner Rammer, TUM
- Kristin H. Braziunas, TUM
- Paola Mairota, UNIBA
- Josef Bruna, ibot cas
- Heike Lischke, WSL
- Daniel Scherrer, nWSL
- Simon Scheiter, senckenberg
- Thomas Hickler, senckenberg
- Tim Anders, senckenberg
- Jessica Hetzer, senckenberg
- Harald Bugmann, Forest Ecology, ETH Zurich
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

```{r, echo=FALSE}
  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=FALSE)
  library(bibtex)
  #knitr::opts_chunk$set(echo = TRUE, fig.pos = "h", out.extra = "")

@@

@@

*Regeneration is a key process in forest dynamics. Models of forest
dynamics have been long lagging an assessment in their capacities to
properly represent this process. Using 13 models of forest dynamics at
the stand, landscape and global levels, we simulate the regeneration
across 200 sites representing environmental gradients across Europe. The
outputs were compared with regeneration data collected from forest
reserves across Europe. In this study we wanted to asses how well are
models of forest dynamics capturing regeneration levels and what can be
changed to improve this. The results show that most of the models
over-estimate regeneration values. However most of them got right trends
in regeneration across temperature and precipitation gradients. Stand
species richness was better capture than recruitment species richness.
Mortality*

@@

Tree recruitment is fundamental for forest dynamics, in particular with
increasing disturbance under climate change and it is also crucial
understand the forest resilience under climate change.

In comparison to recruitment, growth and mortality responses to climate
change are much more studied

Tree recruitment depends on multiple processes (reproduction, fruit
maturation, seed production, dispersal, germination, juvenile growth and
survival) that are all difficult to assess

Models are needed to (1) synthesize existing empirical data and explore
their consequences, and (2) assess future tree recruitment e.g. in the
context of global change scenarios

Regeneration processes in dynamic forest models and how they have been
handled to date [@konig_tree_2022; @bugmann2022].

Tendency of models to be too liberal with recruitment; consequences of
over-recruitment Species richness in regeneration, mortality and
regeneration levels may have strong impacts on forest dynamics

An analysis how well models of forest dynamics perform re. recruitment
is lacking. This is the first time, a large number of forest models
commonly used to assess forest dynamics under climate change is
evaluated against a continental-scale multi-species harmonized data set
on tree recruitment. Briefly praise the dataset [@käber].

In this study wanted to answer the following questions: 1) Are models of
forests dynamics capturing correctly regeneration levels across
environmental gradients 2) Are models of forest dynamics representing
correctly initial composition of tree populations 3) Can models of
forest dynamics correctly capture the initial mortality in tree
in-growth between 7 and 10 cm

@@

A total of 15 models and model variants of forest dynamics at the stand,
landscape and global level were used to simulate stand dynamics and
sample regeneration in 200 sites selected across Europe.

About the models The models ( Table @ref(tab:tableModels)) were created
with different philosophies to represent regeneration. Three models,
LPJ-GUESS, the two variants of ForClim and Treemig simulated more
species than the one asked because XXXX , these species were assigned to
"other" category in the case of TreeMig and in the case of LPJ-GUESS
they were removed from the outputs as they are shrubs species (Boreal
evergreen shrub, Juniperus oxycedrus and quercus coccifera) LPJ-GUESS
also simulate Corylus avellana but only had 3 trees across all samples
and they were removed because due to the low degrees of freedom it was
not possible to create a model for the category "others" for this
model). Forclim simulated a total of 17 species more than the species
asked in the protocol . Observed data also have a category others.

About the sites The sites represent the available range of environmental
gradients in temperature and precipitation (Figure using the variables
degree day and water balance. The sites were selected across a network
of plots that collect information on forest reserves
[@kaber_sheltered_nodate]

Each modeler group was provided with detailed information on how to
perform the simulations including the input variables such as soil or
climate, simulation setup and expected output variables. The overarching
goal of this simulations was the assessment of the regeneration niche as
it arises from empirical data analysis vs. its representation in a wide
range of models of forest dynamics.

There are 35 sites that had a diameter threshold of 10 cm and therefore
we do not know the actual recruitment rates for those sites at 7 cm; yet
the models simulated recruitment for the 7 cm threshold on those sites.
The 35 sites were removed from the simulated data as well to make a fair
comparison.

How many species were simulated per model? why?

Need to say something about zero inflation in the observations vs. the
simulations (current Figure S5)

A GAM model based on XXX was used to compare the trend on the observed
data and the simulated outputs. DESCRIBE GAM

Species simulated: 4C only simulated 4 species. Treemig. ForClim and
Observed data have more species assigned all together into category
"others". LPJ-GUESS simulated bushes that were not consider in the
output of speces composition neither recruitment.

In table @ref(tab:tableModels)

```{r tableModels, echo=FALSE,  message=FALSE}
library(knitr)

modelDF <- readr::read_csv(here::here("data", "modelsDescription.csv"))

modelDF$Type[modelDF$Type ==  "Stand"] <- "S"
modelDF$Type[modelDF$Type ==  "Landscape"] <- "L"
modelDF$Type[modelDF$Type ==  "Global"] <- "G"

modelDF$`Species simulated`[modelDF$`Species simulated` == "F. sylbatica, P. sylvestris, P. abies, A. alba"] <- "F. sylbatica,  P. sylvestris, P. abies,  A. alba"
modelDF$Sampling <- NULL

modelDF$Reference <- "reference"

kable(modelDF,  format = "markdown", caption = "List of models included.")
```

ToDo add a map with the species composition of the observed data

ToDo check in Haralds paper which models are in which category in the
classification of establishment and link this with the results we see
here in structure and complexity. For example iLand and Picus have a
similar structure but results are very different. Are we meeting the
expectations in behaviour fromthe different models?

Species composition in regeneration

We evaluated the species compositions across models and in relation to
the observed data by calculating the Shannon index that defines the
relative frequency of species referring to basal area (BA) (Equation
@ref(eq:shannon))

```{=tex}
\begin{equation}
H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{i})}(\#eq:shannon)
\end{equation}
```
```{=tex}
\begin{equation}
p_{n,i} = \frac{BA_{r_{n,i}}}{BA_{total_{n}}}(\#eq:piShannon)
\end{equation}
```
where, $s$ is the total number of species in the sample $n$ (11 for the
models that simulated all the species and a varied number for those
models that did not consider all the species (see Table
@ref(tab:tableModels) for more details in the simulated species);
$p_{i}$ is the proportion of the species $i$ calculated as the recruited
basal area ($BA_{r_{n,i}}$) for that species $i$ over the total basal
area ($BA_{total_{n}}$) of that sample $n$.

Main hypothesis:

The recruitment species richness is driven by the simulated adult trees
species richness and this is defined on how models represent individual
species recruitment per site with respect of adult proportion

Environmental gradients have an impact on the recruitment of the 5 main
tree species but extremes are not well captured

Models can capture the relative abundance of recruitment of the 5 main
European tree species but they are not doing a good job for minor
species

We evaluate the species composition in the recruitment and for all the
trees present on the site using the relative frequency of species
referring to basal area with the Shannon index calculated per sample,
site and model. The higher the value, the higher the diversity of
species in a particular site and sample.

Mortality in recruitment

Main hypothesis:

the higher regeneration at threshold of 7cm the stronger is the
regeneration mortality, so we will see a larger decease in stem number
between recruitment at threshold of 7 cm and 10 cm (as ratio nn7/nn10
(colname nn710))  In othwr words the number of individuals would
decrease (Reineke self-thinning rule: This is usually calculated from
even age, single species stands. The models could also include the
effects of canopy shading and species competition)

Species composition it is also reduced between 7 and 10 cm

Comments on the Reineke rule

the number of individuals would decrease (Reineke self-thinning rule)
from 7 to 10cm

assumes a relationship between the number of stems and the quadratic
mean diameter in fully stocked pure stands.

This rule is extensively used for management purposes, but it has been
initially calibrated for pure, even-aged stands for relatively few
species.

This is usually calculated from even age, single species stands. The
models could also include the effects of canopy shading and species
competition.

Explanations:

$$Reineke exp rk=-1.605: (7/10)^{(-1.605)} = 1.77$$

# Results

@@

```{r over, echo=FALSE, fig.cap="Mean recruitment values across all samples for the 200 sites. The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold.", out.width='100%'}
        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
```

When looking at the representation of recruitment rates along key
gradients of light availability (basal area), temperature (degree-days)
and soil moisture (water balance), distinct features emerge: the models
reproduce the effect of basal area much better than the effects of
climatic gradients, where they feature varying effects (Figure
\@ref(fig:nvTrend710BA) and FigureS 1-4).

@@

Models did not match the observed patterns in recruitment across soil
moisture and temperature gradients and there was either a clear common
pattern across models. Looking at the soil moisture gradients (FigureS
2), we can see that a group of models decreases recruitment with
increasing soil moisture (4C, ForClim 1, ForClim11, SIBYLA, xComp, PICUS
and LandClim), while few models had the opposite trend (ForCEEPs(f),
FORMIND and LPJ-GUESS). Across temperature gradients all the models that
had a decrease in recruitment with increasing soil moisture showed the
opposite trend (more recruitment with increasing temperature) except for
4C (FigureS 3). The rest of the models did not have a clear pattern
across the temperature gradient.

@@

Most of the models match the level of diversity (as measured by the
Shannon index) of the observed data quite well (Figure \@ref(fig:H710)),
although they consistently underestimate its variability across sites.
Four models overestimate recruitment diversity:4C, PICUS, LANDIS-II and
LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The model
4C is a special case as this model simulated 5 species only. In most
models there are small differences across sites regarding the species
richness of recruitment between 7 and 10 cm. Most models maintain the
recruitment diversity between the 7 and 10 cm thresholds, and the same
is visible from the observed data, there are only few models where the
recruitment diversity increases between 7 and 10 cm which might be an
artifact of the sample strategy and calculating the Shannon index based
in BA as for a same sample recruitment at 10cm will include larger trees
with larger BA and if those are from different species, the Shannon
index will be larger.

@@

Most of the models are capturing well or over-predicting the species
richness at the stand level (Figure \@ref(fig:combinedRichness) A2 and
FigureS 6). The models over predicting species richness at the stand
level are ForClim1, ForClim 11, FORMIND and TreeMig (Figure
\@ref(fig:combinedRichness) A1). However only few models under-predict
the species richness at the stand level (Figure
\@ref(fig:combinedRichness) A3) LPJ-GUESS, ForCEEPS (f), xComp, iLand
and 4C. Species diversity is lower in the recruitment compared to the
stand level, for both observed and simulated data, and for both
recruitment thresholds, i.e. 7 and 10 cm (Figure
\@ref(fig:combinedRichness)B, and FigureS 7). The majority of the models
capture reasonably well the recruitment diversity at lower values of
diversity, however most of them strongly underestimate species diversity
in recruitment for higher observed diversity levels. Models simulating
higher diversity at the stand level do not feature higher diversity in
recruitment (FigureS 6 and 7). Both models with feedback and without
feedback from the adult trees (via seed production) underestimate
species richness in recruitment.

```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site and for observed and simulated values. A) Shows the three trends present across models for the species richness at the stand level. B) Shows the three trends present across models for recruitment at 7cm threshold. The selected representative model per trend group was based on Figure XX and tables available in the supplementary information", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```

>>>>

---
title: "Drivers of tree regeneration remain challenging to be captured by models of forest dynamics"
author: 
- Olalla Díaz-Yáñez, 
- Yannek Käber, 
- Mats Mahnken, 
- Samuel Fischer, 
- Xavier Morin, 
- Manfred Lexer, 
- Christian Hochauer, 
- Katarina Merganicova, 
- Tobias Mette, 
- Werner Rammer, 
- Kristin H. Braziunas, 
- Paola Mairota, 
- Josef Bruna,
- Heike Lischke, 
- Daniel Scherrer, 
- Simon Scheiter, 
- Thomas Hickler, 
- Tim Anders,
- Jessica Hetzer, 
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

```{r, echo=FALSE}
  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries = FALSE)
  library(bibtex)
  #knitr::opts_chunk$set(echo = TRUE, fig.pos = "h", out.extra = "")

@@

    Word count: `r as.integer(sub("(\\d+).+$", "\\1", system(sprintf("wc -w %s", knitr::current_input()), intern = TRUE))) - 20`

@@

Regeneration is a key process in forest dynamics. Models of forest
dynamics have been long lagging an assessment in their capacities to
properly represent this process. Using 15 models and variants of forest
dynamics at the stand, landscape and global levels, we simulate the
regeneration across 200 sites representing environmental gradients
across Europe. The simulation results were compared with observed data
collected from forest reserves across Europe. In this study we want to
asses how well are models of forest dynamics capturing regeneration
levels, species composition and mortality in regeneration. The results
show that .

@@

Forests provide multiple products and services of vital importance.
Under the pressure of climate change and increasing disturbances, it is
even more important that we understand how forests will change and
evaluate their capacity to adapt or be resilient to these changes
[@lindner2010]. A wide range of models of forest dynamics were developed
over the past decades considering the impacts of climate in forest
dynamics [@bugmann2022]. From these studies, it is becoming increasingly
evident that while we have a reasonably good understanding how to model
tree growth [@vanclay1997; @bugmann1996], and substantial efforts have
been dedicated to improving the depiction of tree mortality processes
[@bugmann2019], tree regeneration remains a less studied process in most
of these models, while becoming ever more important in the context of
forest disturbances and forest resilience.

Tree recruitment depends on multiple processes such as reproduction,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival; these processes are all difficult to assess and
some of them are scarcely understood and thus appear highly stochastic.
Finding the right level of complexity in the definition of the factors
leading to tree recruitment is challenging and models of forest dynamics
have used different approaches [@bugmann2022]. Overall, models are
needed to (1) synthesize existing empirical data and explore their
consequences, and (2) assess future tree recruitment e.g. in the context
of global change scenarios. Currently, regeneration processes in dynamic
forest models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022], from entirely ignoring tree regeneration (as done in
classical forest growth models, e.g. @pretzsch2002), across the use of a
few simple environmental filters, as done in most forest \"gap\" models
(e.g., @shuman2011) and Dynamic Global Vegetation Models (e.g.,
@smith2001; @hickler2012), to complex approaches that incorporate local
feedback from the canopy, multiple ecological processes and often also
small time steps (e.g., @wehrli2006; @seidl2012), or field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time (e.g., @ribbens1994).

With the current strategies that models of forest dynamics are using to
represent tree recruitment we are obtaining results that faced several
problems related to over-recruitment, needed excess of mortality at
early stages and inadecuate species composition in regeneration which
have strong impacts on forest dynamics. Some model outputs are
over-recruiting which subsequently must die primarily by competition in
later stages. This is characteristic of \'classical\' forest gap models
[@gutiérrez2016], at least as long as the simulations comprise a
multitude of species. Other models use a model output to correct
recruitment levels for example by routinely calibrate them against the
predicted canopy species composition affecting as well the species
dominance.

Correctly capturing the species composition in recruitment is often also
a challenge, thus leading to unrealistic successional drift in the model
REFERENCE?, for example by the change of the dominant species via
stochastic processes that continuously bring seeds of other species that
the come to dominance, with this strategy a new species might start to
rise eliminating the former dominant species. A related issues is the
rarification of species richness due to positive feedback effects such
that eventually just single-species forest stands remain. This is
sometimes handled by the incorporation of a low level of seed influx of
all species at all times (e.g., @lischke2006 ). However, simulated
species composition is usually exceedingly sensitive to that level
whereas the parameters of such functions are poorly constrained by field
data. There is also often a problem with the insuficient constraint of
parameter values for models that start from very small tree sizes (e.g.,
10 cm of height or even seeds) and track tree development in a
process-oriented manner by considering a multitude of ecological
influences, rather than being statistically driven and emphasizing
recruitment into a larger size class (e.g., @wehrli2006 ). Most of these
undesired behaviors in tree recruitment are often not published as such,
but in the vast majority of studies they are corrected (e.g.,
@rasche2011 , where for most study sites regeneration was simply
switched off). Thus, a strong focus on the modeling of tree regeneration
processes is highly important and long overdue [@price] if we are to
make reliable projections of future forest dynamics, i.e. when the
models need to be operated in extrapolation mode, but also from a
fundamental ecological point of view (increasing systems understanding).

This is the first time, a large number of forest models commonly used to
assess forest dynamics under climate change are evaluated against a
continental-scale multi-species harmonized data set on tree recruitment
[@käber2022]. Tree recruitment is understudied in field studies and
modelling [@hanburybrown_forest_2022], because little data has been
available and data was very fragmented, i.e., not in one harmonized data
set. In this study we relied on models with various 'philosophies',
including models operating at the stand to the global scale. Because of
the large variability in recruitment patterns in nature and the large
amount of factor driving recruitment, including some which are not
included in most models, such as deer browsing, we do not aim at a
detailed statistical evaluation of each model. Instead, we aim at
evaluating general recruitment patterns, magnitudes and benchmark the
regeneration niche of multiple species based on empirical data along a
wide environmental gradient (temperature, moisture, and light
availability). More specifically in this study we wanted to answer the
following questions: 1) are models of forests dynamics capturing
correctly regeneration levels across environmental gradients; 2) are
models of forest dynamics representing correctly initial composition of
tree populations;and, 3) can models of forest dynamics correctly capture
the initial mortality in tree in-growth?.

@@

## Models

Fifteen models of forest dynamics (including two model variants) were
used to assess tree regeneration. The models selected simulate forest
dynamics at the stand, landscape and global scale (Table
\@ref(tab:tableModels)). Models construction philosophies and origin
differs, most of the models have a largely process based approach and
two models have formulations derived from statistical analyses based on
inventory data (SYBILA and xComp). Some of the models largely rely on
the original approach underlying forest gap models (ForClim 1, ForClim
11, LandClim, Landis II ); and some include plant ecophysiology
formulations (FORMIND, PICUS, 4C, iLand). This different approaches in
the models construction also has influenced how tree establishment is
defined.

Based on their establishment modules the models can be categorized as
regeneration or recruitment (or ingrowth) models [@vanclay1994;
@konig_tree_2022] (see Table \@ref(tab:tableModels)). Regeneration
models include processes such as flowering and pollination, seed
production, seed dispersal, germination or seedling establishment that
lead to the number of seedlings or saplings. Recruitment models
introduce a number of seedlings or saplings with certain characteristic
such as biomass or dbh, without considering earlier development
processes. We can also distinguish models that have feedback in the
regeneration as those in which the level and species composition in
regeneration is influenced by the existing adult tree community via the
production of seeds, seedlings or saplings. The starting point for
regeneration it is also different per model varying from seedbank, seed,
sapling or ingrowth. The overall complexity in the establishment modules
also varies, following @bugmann2022 we can see that most of the models
included in this study can be classified as having rather low complexity
in their tree recruitment modules while few of them have high complexity
specifically when defining processes such as XXXX, XXXX (enumerate
models). Further details on the models specific formulation related to
their regeneration modules and an overall summary across models is
presented in the supplementary materials (SHALL WE DO THIS?).

```{r tableModels, echo=FALSE,  message=FALSE, warning=FALSE}
library(knitr)

modelDF <- readr::read_csv(here::here("data", "modelsDescription.csv"))
modelDF$`Species simulated`[modelDF$`Species simulated` == "F. sylbatica, P. sylvestris, P. abies, A. alba"] <- "F. sylbatica,  P. sylvestris, P. abies,  A. alba"
modelDF$Sampling <- NULL

kable(modelDF, 
      #format = "markdown",
      format = "latex", booktabs = TRUE,
      caption = "List of models included.") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |> 
      kableExtra::column_spec(column = 2, width = "5em") |> 
      kableExtra::column_spec(column = 3, width = "5em") |> 
      kableExtra::column_spec(column = 4, width = "5em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "5em") |>
      kableExtra::column_spec(column = 7, width = "8em") |>
      kableExtra::column_spec(column = 8, width = "5em") |>
      kableExtra::column_spec(column = 9, width = "5em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "5em") |>
      kableExtra::column_spec(column = 12, width = "5em") 
```

## Observed data

Regeneration data covering a wide range of environmental gradients is
hard to obtain and this is one of the reasons why most of the models of
forest dynamics have never been confronted with a data-set covering a
wide range of climate gradients in a large number sites to evaluate how
well are they capturing regeneration. The observed regeneration data
used in this study represent the available range of environmental
gradients in temperature and precipitation in Europe using 200 sites of
strict forest reserves compiled in the framework of the EuFoRIa network
(Figure \@ref(fig:map)). The sample forests have not been managed in at
least 10 years (most of them in close to 100 years?), and each of the
sites contains at least two consecutive measurements. Regeneration was
measured at 7 and 10 cm threshold in all the sites except for 35 sites
where only regeneration at 10 cm diameter threshold was available. For
more details on this dataset and their regeneration information please
refer to @käber2022.

## Simulation protocol

Each modeler group was provided with detailed protocol (available in the
supplementary material) with instruction on how to perform the
simulations regarding input variables such as climate or soil,
simulation setup and expected output variables. No information was
provided on the regeneration prior to the simulation. The overarching
goal of this experiment was the assessment of the regeneration niche as
it arises from empirical data analysis against its representation in a
wide range of models of forest dynamics.

The input variables were collected and aggregated to be adapted to the
different model demands. The time series of climatic variables from 1981
to 2018 were provided in hourly (Era5-land data, @muñozsabater2019),
daily and monthly resolution (CHELSA data, @karger2021). Some variables
required for some models such as relative humidity or vapor pressure
deficit had to be calculated from these variables. The final strategy on
how to use the climate data was different in each model based on the
approach the better fitted the model (Table \@ref(tab:tableModels)). The
soil quality data was provided as a continuous value between 1 and 5
(Soilgrids dataset, @hengl2017, @poggio2021). The protocol also provided
the elevation, slope and aspect for each of the 200 sites
[@asterscienceteam2019], no other spatial information such as
coordinates were given with the exception of iLand that used the blurred
coordinates to obtain detailed soil information values for the different
sites.

The simulations had to be run in the absence of natural disturbances,
one the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE [ONE
SENTENCE]{.underline} OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE.
The simulations had to be set up in a way that they will obtain samples
on species specific regeneration rates per decade and per ha. The
modeling teams decided on the simulated area and how they derive these
samples. Further details on how each modeling team prepared the
simulations and the outputs are available in the supplementary material.

The simulations were run from bare ground in the absence of management
to a simulated equilibrium ("potential natural vegetation") with the
current climate as the ultimate goal was to derive the regeneration
niche. The regeneration niche is defined as the passing of a diameter
threshold of 7 and 10 cm (ingrowth). Both the width of the regeneration
niche (i.e., in environmental space) as well as the intensity of the
recruitment process (i.e., the number of ingrowth trees per area and per
unit of time) were of interest. The exact length of the simulation was
decided by the modeling teams (Table \@ref(tab:tableModels)).

The simulations were run for mixed species forests (not multiple
single-species simulations) using mixtures of the following species or
genera: *Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus,
Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior,
Quercus spp., Alnus glutinosa, Pinus sylvestris.* Three models included
additional species in the simulations (ForClim 1, ForClim 11 and
TreeMig) and three could not simulate all the species (4C, xComp,
TreeMig, LPJ-GUESS) (Table \@ref(tab:tableModels)). HERE **FORCLIM**
PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED MORE SPECIES.
HERE **TREEMIG** PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED
MORE SPECIES AND MISSED ONE. HERE **4C** PROVIDES [ONE
SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. HERE **XCOMP**
PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. The
model aDGVM2 did not have species level option and provided the default
approach with HERE **aDVGM2** PROVIDES [ONE SENTENCE]{.underline} WITH
THE INFORMATION ON THE DEFAULT SET UP FOR THE SPECIES.

Each of the models reported the number of recruits over a 10 year
interval per species and per ha for each of the sites. To better
understand the variation per site each modeling team decided on how to
sample 200 multiple values per site to show the variation in recruitment
for spatial units of 1 ha during the simulated phase of potential
natural vegetation but without providing average of multiple 1-ha
samples. This was done with different strategies such as sampling
simulated data from the same 1-ha plot in the equilibrium over time, or
by sampling several 1-ha plots from the simulated forest at one specific
point in time (in the equilibrium). This resulted in 880000 observation
per model that simulated the 11 species indicated in the protocol (200
sites, 200 samples per site, 11 species and two diameter thresholds).
Those models that simulated extra species, the recruitment of those
extra species were aggregated in a category others. The simulated data
per model and the observations used in this study are provided as a
supplementary information OR IN A REPOSITORY? available here(XXX).

```{r map, echo=FALSE, fig.cap="The 200 sites with the dominant tree species in the observed recruitment at 10 cm threshold.", out.width='80%'}
knitr::include_graphics(here::here("figures", "theMap.png"))
```

## Analysis

The simulation results were analyzed in regard of their regeneration
levels, species composition and mortality in regeneration. The observed
data was modeled using a generalized additive model with a negative
binomial distribution to better understand the trend and regeneration
levels of the observed regeneration and the simulation results per
model. The regeneration levels were evaluated against the stand basal
area (the basal area of all the trees in each sample) as a proxy of
light availability for regeneration, the degree-days as a proxy for
temperature and the water balance as a proxy for soil moisture. The mean
recruited values were calculated as the mean across samples per site and
when the regeneration levels were evaluated across the total basal area
and climate gradients the sites.

We evaluated the species composition across models and in relation to
the observed data by calculating the Shannon index that defines the
relative frequency of species referring to basal area where the higher
the value, the higher the diversity of species in a particular site and
sample (Equation \ref{eqpiShannon}).

```{=tex}
\begin{equation}
H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{i})}\label{eqpiShannon}
\end{equation}
```
```{=tex}
\begin{equation}
p_{n,i} = \frac{r_{BA_{n,i}}}{Total_{BA_{n}}} \label{eqpiShannon2}
\end{equation}
```
where, $s$ is the total number of species simulated $n$ (11 for the
models that simulated all the species and a varied number for those
models that did not consider all the species, see Table
\@ref(tab:tableModels) for more details in the simulated species);
$p_{i}$ is the proportion of the species $i$ calculated as the recruited
basal area ($r_{BA_{n,i}}$) for that species $i$ over the stand basal
area ($Total_{BA_{n}}$) of the sample $n$.

In order to understand the impact of adult species composition on
recruitment across climatic gradients we focused on the five main tree
species: *Abies alba*, *Fagus sylvatica*, *Picea abies*, *Pinus
sylvestris*, and *Quercus* spp. For these species we calculated the mean
recruited basal area share per site as the mean value across the
available samples per site:

```{=tex}
\begin{equation}
\bar{R}_{BA_{share_{t,i}}} = \frac{\bar{r}_{t,i}}{\sum_{i = 1}^{s}{\bar{r}_{t,i}}} \label{eq:BAshare}
\end{equation}
```
where, $s$ is the total number of species simulated; $\bar{r}_{n,i}$ is
the mean recruitment of the species $i$ per site $t$ across the
available samples on site.

There were 35 sites that had a diameter threshold of 10 cm and therefore
we do not know the actual recruitment rates for those sites at 7 cm; yet
the models simulated recruitment for the 7 cm threshold on those sites.
The 35 sites were removed from the simulated data to make a fair
comparison. Furthermore, when looking at the recruited basal area share
per species ($\bar{R}_{BA_{share_{i}}}$) we categorized the basal area
share as zero if the total recruited basal area (
$\sum_{i = 1}^{s}{\bar{r}_{t,i}}$ ) equal zero and the recruited basal
area of that species ($\bar{r}_{t,i}$) was also zero.

Mortality at regeneration was assessed based on the difference on
recruitment between 7 and 10 cm diameter thresholds. We used the Reineke
self-thinning rule [@reineke1933; @pretzsch2003] as a reference to
estimate if the different recruitment values were above or bellow the
expected change between these two diameters. The Reineke self-thinning
rule is usually calculated from even-aged, single species stands and
assumes a relationship between the number of stems and the quadratic
mean diameter in fully stocked pure stands. The value used in our
comparisons was 1.77 and was calculated following the Equation
\ref{eqReineke}.

```{=tex}
\begin{equation}
Reineke = (7/10)^{(-1.605)} \label{eqReineke}
\end{equation}
```
# Results

@@

```{r over, echo=FALSE, fig.cap="Mean recruitment values across all samples for the 200 sites. The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
```

When looking at the representation of recruitment rates along key
gradients of light availability (basal area), temperature (degree-days)
and soil moisture (water balance), distinct features emerge: the models
reproduce the effect of basal area much better than the effects of
climatic gradients, where they feature varying effects (Figure
\@ref(fig:nvTrend710BA) and supplementary figures 1 and 2).

@@

Models did not match the observed patterns in recruitment across soil
moisture and temperature gradients and there was either a clear common
pattern across models. Looking at the soil moisture gradients
(supplementary figure 1), we can see that a group of models decreases
recruitment with increasing soil moisture (4C, ForClim 1, ForClim11,
SIBYLA, xComp, PICUS and LandClim), while few models had the opposite
trend (ForCEEPs(f), FORMIND and LPJ-GUESS). Across temperature gradients
all the models that had a decrease in recruitment with increasing soil
moisture showed the opposite trend (more recruitment with increasing
temperature) except for 4C (supplementary figure 2). The rest of the
models did not have a clear pattern across the temperature gradient.

Both observations and simulated data had no recruitment in some samples
and sites (supplementary table 1). The observed data had over 3% of the
samples with no recruitment. Only three models had larger proportion of
no recruitment (4C, Landis II and aDVM2). Three models always simulate
recruitment for both 7 and 10 cm threshold (TreeMig and LPJ-GUESS), an
the rest of the models have a very low percentage of samples with no
recruitment (0.01% - 1.31%).

@@

Most of the models match the level of diversity (as measured by the
Shannon index) of the observed data quite well (Figure \@ref(fig:H710)),
although they consistently underestimate its variability across sites.
Four models overestimate recruitment diversity: 4C, PICUS, LANDIS-II and
LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The model
4C is a special case as this model simulated 5 species only. In most
models there are small differences across sites regarding the species
richness of recruitment between 7 and 10 cm. Most models maintain the
recruitment diversity between the 7 and 10 cm thresholds, and the same
is visible from the observed data, there are only few models where the
recruitment diversity increases between 7 and 10 cm which might be an
artifact of the sample strategy and calculating the Shannon index based
in basal area as for a same sample recruitment at 10cm will include
larger trees with larger basal area and if those are from different
species, the Shannon index will be larger.

@@

Species diversity is lower in the recruitment compared to the stand
level, for both observed and simulated data, and for both recruitment
thresholds, i.e. 7 and 10 cm (supplementary figure 3 and 4). Models
simulating higher diversity at the stand level do not feature higher
diversity in recruitment (supplementary figure 3 and 4). Most of the
models are capturing well or over-predicting the species richness at the
stand level (Figure \@ref(fig:combinedRichness)-A-centered plot and
supplementary figure 5). The models over predicting species richness at
the stand level are ForClim1, ForClim 11, FORMIND and TreeMig (Figure
\@ref(fig:combinedRichness)-A-left plot). However only few models
under-predict the species richness at the stand level LPJ-GUESS,
ForCEEPS (f), xComp, iLand and 4C. (Figure
\@ref(fig:combinedRichness)-A-right plot). The majority of the models
capture reasonably well the recruitment diversity at lower values of
diversity, however most of them strongly underestimate species diversity
in recruitment for higher observed diversity levels(Figure
\@ref(fig:combinedRichness)-B and supplementary figure 6). Both models
with feedback and without feedback from the adult trees (via seed
production) underestimate species richness in recruitment compared to
the observed data.

```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site and for observed and simulated values. A) Shows the three trends present across models for the species richness at the stand level. B) Shows the three trends present across models for recruitment at 7cm threshold. The selected representative model per trend group was based on supplementary figure 5 and 6", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```
